# (PART\*) Student Guide {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Set Up

## Set Up Your Computer

In order to run your analyses, you will use the [AnVIL cloud computing platform](https://anvilproject.org/), so that you do not need to install everything on your own computer.  AnVIL (Analysis Visualization and Informatics Lab-space) is a cloud computing platform specially designed for analyzing biological data, and is used by scientists doing all sorts of biological research.

**Create Google Account**

Behind the scenes, AnVIL relies on Google Cloud Platform to provide computing infrastructure, so the first thing you will need to do is set up (free) a Google account.

If you do not already have a Google account that you would like to use for accessing AnVIL, [create one now](https://accounts.google.com/SignUp).


# Introduction

# Exercises

## Complete your first `swirl` lesson

*Estimated time: 20 min*

1. Open RStudio.
1. In the R console window, type the following commands to load the swirl package:

    ```
    library(swirl)
    swirl()
    ```

1. Install the course “**R Programming: The basics of programming in R**”, by following the instructions provided by swirl.
    a. Enter your name.
    a. Press `ENTER`.
    a. Enter `1`, `2`, or `3`.
    a. Install the course “R Programming: The basics of programming in R”.
1. Complete your first lesson.
    a. Select “**R Programming**”.
    a. Select “**Basic Building Blocks**”, by entering `1` ( i.e. the corresponding lesson number ).
    a. Follow the instructions provided by swirl to complete your first lesson.
    a. At the end, when swirl asks if you would like to receive Coursera credit, select “No”.


## Complete additional `swirl` lessons

*Estimated time: 80 min*

1. Complete additional swirl lessons. 
    a. Select “**R Programming**”
    a. Compete each of the following lessons:
        - **4: Vectors**
        - **5: Missing_Values**
        - **6: Subsetting_Vectors**
        - **12: Looking_at_Data**
1. Once you have completed all the lessons listed above, exit swirl.
    a. To exit swirl, type `bye()` in the R console, press `ESC` on your keyboard, or enter `0` in response to the swirl course menu prompt.

# Project: Explore dataframe

Now we will use your new R skills to explore a data.frame!

<br>

## R tips

Don't worry if you can't remember everything!  Data scientists and computer programmers don't memorize everything either.  That's why there's a help function built in to R!

As a reminder, here are a few useful commands for exploring dataframes:
```
dim()
head()
tail()
summary()
str()
```

If you need help remembering how to use a command, use `?`
```
?dim
```

If you start using R frequently, you'll naturally start to remember some of the common commands.  But don't hesitate to look things up, or ask if you get stuck!

## Explore the ??? dataframe

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/dataset/awesomedata/_child_explore_awesomedata.Rmd",
  repo_name = "jhudsl/GDSCN_Book_swirl"
)
```

# Wrap-up

Cloud computing costs are based on the amount of time you use the computing resources, so it's important to clean up after yourself when you're done, and not just leave the computers running.

There are two ways to "shut down" RStudio on AnVIL:

- **Pause** the environment: This will save a copy of your work, and then release the computers for other people to use them.  **Do this if you plan to continue working in RStudio**.
    - It's similar to turning off your computer or phone - when you start it back up, everything will be where you left it.
    - This still costs a small amount of money, but much less than leaving the computer running.  
- **Delete** the environment: This will delete everything and then release the computers for other people to use them.  **Do this if you are completely finished working in RStudio**, or if your future work will be in a new environment.
    - It's similar to throwing your computer or phone in the trash!
    - **You will not be able to recover your work.**
    - Make sure you have saved anything you need before you delete your environment.  

## Pause RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_pause.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

## Delete RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_delete.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
