# (PART\*) Student Guide {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Setup

# Introduction

# Exercises

## Complete your first `swirl` lesson

*Estimated time: 20 min*

1. Open RStudio.
1. In the R console window, type the following commands to load the swirl package:

    ```
    library(swirl)
    swirl()
    ```

1. Install the course “**R Programming: The basics of programming in R**”, by following the instructions provided by swirl.
    a. Enter your name.
    a. Press `ENTER`.
    a. Enter `1`, `2`, or `3`.
    a. Install the course “R Programming: The basics of programming in R”.
1. Complete your first lesson.
    a. Select “**R Programming**”.
    a. Select “**Basic Building Blocks**”, by entering `1` ( i.e. the corresponding lesson number ).
    a. Follow the instructions provided by swirl to complete your first lesson.
    a. At the end, when swirl asks if you would like to receive Coursera credit, select “No”.


## Complete additional `swirl` lessons

*Estimated time: 80 min*

1. Complete additional swirl lessons. 
    a. Select “**R Programming**”
    a. Compete each of the following lessons:
        - **4: Vectors**
        - **5: Missing_Values**
        - **6: Subsetting_Vectors**
        - **12: Looking_at_Data**
1. Once you have completed all the lessons listed above, exit swirl.
    a. To exit swirl, type `bye()` in the R console, press `ESC` on your keyboard, or enter `0` in response to the swirl course menu prompt.

# Project: Explore dataframe

Now we will use your new R skills to explore a data.frame.

<br>

## R tips

Don't worry if you can't remember everything - data scientists and computer programmers don't memorize everything either.  That's why there's a help function built in to R!

As a reminder, here are a few useful commands for exploring dataframes:
```
dim()
head()
tail()
summary()
str()
```

If you need help remembering how to use a command, use `?`
```
?dim
```

If you start using R frequently, you'll naturally start to remember some of the common commands.  But don't hesitate to look things up, or ask if you get stuck!

## Explore dataframe

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/dataset/us_contagious_diseases/_child_explore_us_contagious_diseases.Rmd",
  branch = "dslabs-contagious-disease",
  repo_name = "jhudsl/GDSCN_Book_swirl"
)
```

To start, you will need to load the `dslabs` package, which contains the dataset.  Type the following commands to load the package and confirm that you can view the dataset.  You should see a printout of the first few rows of data, as shown below.

```{r}
library(dslabs)
head(us_contagious_diseases)
```
 
:::{.notice}
It can be a bit annoying to have to type out the whole dataset name, and it's easy to make errors.  Remember that you can use "tab completion" to make the computer do the work for you!

- Start typing the first few letters (i.e. `us`), then hit the `TAB` key on your keyboard and you should see a list of autocomplete possibilities.
- Use your mouse or the arrow keys to select the one you want, or just keep typing until it's the only one left, then hit `ENTER` and it will autocomplete for you.
:::

Take a minute to look at the first few rows of this dataset.

:::{.reflection}
1. How many columns are there?
1. What's in each column?
:::

Now use some R commands dig a little deeper into the data.

:::{.reflection}
1. What diseases are reported in this dataset?  (Hint: there are 7)
1. What time period does this dataset cover?  In other words
    a. What's the earliest year reported?
    b. What's the latest year reported?
:::

# Wrap-up

Cloud computing costs are based on the amount of time you use the computing resources, so it's important to clean up after yourself when you're done, and not just leave the computers running.

There are two ways to "shut down" RStudio on AnVIL:

- **Pause** the environment: This will save a copy of your work, and then release the computers for other people to use them.  **Do this if you plan to continue working in RStudio**.
    - It's similar to turning off your computer or phone - when you start it back up, everything will be where you left it.
    - This still costs a small amount of money, but much less than leaving the computer running.  
- **Delete** the environment: This will delete everything and then release the computers for other people to use them.  **Do this if you are completely finished working in RStudio**, or if your future work will be in a new environment.
    - It's similar to throwing your computer or phone in the trash!
    - **You will not be able to recover your work.**
    - Make sure you have saved anything you need before you delete your environment.  

## Pause RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_pause.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

## Delete RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_delete.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
