# (PART\*) Student Guide {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Set Up

In order to run your analyses, you will use the [AnVIL cloud computing platform](https://anvilproject.org/), so that you do not need to install everything on your own computer.

<br>

AnVIL (Analysis Visualization and Informatics Lab-space) is a cloud computing platform specially designed for analyzing biological data, and is used by scientists doing all sorts of biological research.

:::{.notice}
**AnVIL in a nutshell**

- Behind the scenes, AnVIL relies on Google Cloud Platform to provide computing infrastructure.  Basically, AnVIL lets you "rent" computers from Google (remotely).  Whenever you run an analyses on AnVIL, it actually runs on one of Google's computers, and AnVIL lets you see the results in your browser.
- AnVIL uses [Terra](https://anvil.terra.bio/) to provides many computational tools useful for biological data analysis, such as [RStudio](https://www.rstudio.com/products/rstudio/), [Galaxy](https://usegalaxy.org/), and [Jupyter Notebooks](https://jupyter.org/).  Terra takes care of installing these tools on Google's computers, so that you can just start using them.
:::

## Create Google Account

The first thing you will need to do is set up (free) a Google account.

If you do not already have a Google account that you would like to use for accessing AnVIL, [create one now](https://accounts.google.com/SignUp).

- Alternatively, if you would like to create a Google account that is associated an existing non-Gmail email address, you can follow [these instructions](https://support.terra.bio/hc/en-us/articles/360029186611).

## Step 2: Log In to Terra

Next, make sure you can log in to Terra - you will need to use Terra to perform computations on AnVIL. 

You can access Terra by going to [`anvil.terra.bio`](https://anvil.terra.bio/), or by clicking the link on the AnVIL home page.

```{r, echo=FALSE, fig.alt='Screenshot of the AnVIL home page. The button to "launch Terra" is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1tyt8ifxD3y7QW3nN27jh02aPrBvNEo1GN_u2Ne_UngM/edit#slide=id.gde7783227e_0_31")
```

Open Terra, and you should be prompted to sign in with your Google account.

## Step 3: Share username

Finally, make sure your instructor has your Google account username (e.g. `myname@gmail.com`), so they can give you access to everything you need.

- Make sure there are no typos!
- If you have multiple Google accounts, make sure you give them the username that you will be using to access AnVIL

:::{.warning}
It is *very important* that you share the Google account you will be using to access AnVIL with with your instructor! Otherwise, the instructor cannot add you to billing projects or Workspaces, and you will be unable to proceed with your assignments.
:::


# Introduction

# Exercises

## Complete your first `swirl` lesson

*Estimated time: 20 min*

1. Open RStudio.
1. In the R console window, type the following commands to load the swirl package:

    ```
    library(swirl)
    swirl()
    ```

1. Install the course “**R Programming: The basics of programming in R**”, by following the instructions provided by swirl.
    a. Enter your name.
    a. Press `ENTER`.
    a. Enter `1`, `2`, or `3`.
    a. Install the course “R Programming: The basics of programming in R”.
1. Complete your first lesson.
    a. Select “**R Programming**”.
    a. Select “**Basic Building Blocks**”, by entering `1` ( i.e. the corresponding lesson number ).
    a. Follow the instructions provided by swirl to complete your first lesson.
    a. At the end, when swirl asks if you would like to receive Coursera credit, select “No”.


## Complete additional `swirl` lessons

*Estimated time: 80 min*

1. Complete additional swirl lessons. 
    a. Select “**R Programming**”
    a. Compete each of the following lessons:
        - **4: Vectors**
        - **5: Missing_Values**
        - **6: Subsetting_Vectors**
        - **12: Looking_at_Data**
1. Once you have completed all the lessons listed above, exit swirl.
    a. To exit swirl, type `bye()` in the R console, press `ESC` on your keyboard, or enter `0` in response to the swirl course menu prompt.

# Project: Explore dataframe

Now we will use your new R skills to explore a data.frame!

<br>

## R tips

Don't worry if you can't remember everything!  Data scientists and computer programmers don't memorize everything either.  That's why there's a help function built in to R!

As a reminder, here are a few useful commands for exploring dataframes:
```
dim()
head()
tail()
summary()
str()
```

If you need help remembering how to use a command, use `?`
```
?dim
```

If you start using R frequently, you'll naturally start to remember some of the common commands.  But don't hesitate to look things up, or ask if you get stuck!

## Explore the ??? dataframe

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/dataset/awesomedata/_child_explore_awesomedata.Rmd",
  repo_name = "jhudsl/GDSCN_Book_swirl"
)
```

# Wrap-up

Cloud computing costs are based on the amount of time you use the computing resources, so it's important to clean up after yourself when you're done, and not just leave the computers running.

There are two ways to "shut down" RStudio on AnVIL:

- **Pause** the environment: This will save a copy of your work, and then release the computers for other people to use them.  **Do this if you plan to continue working in RStudio**.
    - It's similar to turning off your computer or phone - when you start it back up, everything will be where you left it.
    - This still costs a small amount of money, but much less than leaving the computer running.  
- **Delete** the environment: This will delete everything and then release the computers for other people to use them.  **Do this if you are completely finished working in RStudio**, or if your future work will be in a new environment.
    - It's similar to throwing your computer or phone in the trash!
    - **You will not be able to recover your work.**
    - Make sure you have saved anything you need before you delete your environment.  

## Pause RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_pause.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

## Delete RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_delete.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
