# (PART\*) Student Guide {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Setup

# Introduction

# Exercises

## Complete your first `swirl` lesson

*Estimated time: 20 min*

1. Open RStudio.
1. In the R console window, type the following commands to load the swirl package:

    ```
    library(swirl)
    swirl()
    ```

1. Install the course “**R Programming: The basics of programming in R**”, by following the instructions provided by swirl.
    a. Enter your name.
    a. Press `ENTER`.
    a. Enter `1`, `2`, or `3`.
    a. Install the course “R Programming: The basics of programming in R”.
1. Complete your first lesson.
    a. Select “**R Programming**”.
    a. Select “**Basic Building Blocks**”, by entering `1` ( i.e. the corresponding lesson number ).
    a. Follow the instructions provided by swirl to complete your first lesson.
    a. At the end, when swirl asks if you would like to receive Coursera credit, select “No”.


## Complete additional `swirl` lessons

*Estimated time: 80 min*

1. Complete additional swirl lessons. 
    a. Select “**R Programming**”
    a. Compete each of the following lessons:
        - **4: Vectors**
        - **5: Missing_Values**
        - **6: Subsetting_Vectors**
        - **12: Looking_at_Data**
1. Once you have completed all the lessons listed above, exit swirl.
    a. To exit swirl, type `bye()` in the R console, press `ESC` on your keyboard, or enter `0` in response to the swirl course menu prompt.


# Wrap-up

Cloud computing costs are based on the amount of time you use the computing resources, so it's important to clean up after yourself when you're done, and not just leave the computers running.

There are two ways to "shut down" RStudio on AnVIL:

- **Pause** the environment: This will save a copy of your work, and then release the computers for other people to use them.  **Do this if you plan to continue working in RStudio**.
    - It's similar to turning off your computer or phone - when you start it back up, everything will be where you left it.
    - This still costs a small amount of money, but much less than leaving the computer running.  
- **Delete** the environment: This will delete everything and then release the computers for other people to use them.  **Do this if you are completely finished working in RStudio**, or if your future work will be in a new environment.
    - It's similar to throwing your computer or phone in the trash!
    - **You will not be able to recover your work.**
    - Make sure you have saved anything you need before you delete your environment.  

## Pause RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_pause.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

## Delete RStudio environment

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_delete.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
