# Setup on AnVIL
In order to access R and RStudio on AnVIL, there are a few steps you need to complete.

<br>

:::{.notice}
You may be able to skip some of these setup steps if you did them ahead of time.
:::

## Set up AnVIL Account

```{r, echo = FALSE, results='asis'}
cow::borrow_chapter(
  doc_path = "child/_child_student_create_account.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

## Access Materials (Clone Workspace)

```{r, echo = FALSE, results='asis'}
#workspace_name <- "Example_Workspace"
#workspace_link <- "http://example.com/"
billing_project <- "provided by your instructor"
cow::borrow_chapter(
  doc_path = "child/_child_student_workspace_clone.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```

:::{.warning}
**Pause here** if you are not going to be doing the lab exercises right away.  Once you start up RStudio, it will cost money to keep it running.  It costs a few cents an hour, so it's quite cheap as long as you use it responsibly.  But it can add up if you leave it running for days or weeks when you don't need it.
:::

## Start RStudio

```{r, echo = FALSE, results='asis'}
AnVIL_module_settings <- list(
  audience = "student",
#   docker_image = "example docker image link",
  startup_script = "ask your instructor"
)
cow::borrow_chapter(
  doc_path = "child/_child_rstudio_launch.Rmd",
  repo_name = "jhudsl/AnVIL_Template"
)
```
