
1. Stopping your cloud environment only pauses your work. When you are ready to delete the cloud environment, click on the gear icon in the upper right corner to “Update cloud environment”.

    ```{r, echo=FALSE, fig.alt='Screenshot of the Workspace page. The gear icon on the top right that updates the cloud environment is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1eypYLLqD11-NwHLs4adGpcuSB07dYEJfAaALSMvgzqw/edit#slide=id.ge1182913a6_0_41")
    ```

1. Click on “Delete Environment Options”.

    ```{r, echo=FALSE, fig.alt='Screenshot of the cloud environment popout. "Delete environment options" is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1eypYLLqD11-NwHLs4adGpcuSB07dYEJfAaALSMvgzqw/edit#slide=id.ge1182913a6_0_20")
    ```

1. If you are certain that you do not need the data and configuration on your disk, you should select "Delete everything, including persistent disk".  If there is anything you would like to save, open the compute environment and copy the file(s) from your compute environment to another location, such as the Workspace bucket, GitHub, or your local machine, depending on your needs.

    ```{r, echo=FALSE, fig.alt='Screenshot of the cloud environment popout. "Delete everything, including persistent disk" is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1eypYLLqD11-NwHLs4adGpcuSB07dYEJfAaALSMvgzqw/edit#slide=id.ge1182913a6_0_46")
    ```

1. Select "DELETE".

    ```{r, echo=FALSE, fig.alt='Screenshot of the cloud environment popout. "Delete" is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1eypYLLqD11-NwHLs4adGpcuSB07dYEJfAaALSMvgzqw/edit#slide=id.ge1182913a6_0_51")
    ```
